---
title: "Reproducible Research"
author: "Yao Tian"
date: "February 23, 2017"
output: html_document
---

This is for the project of week 2 "Reproduciable research"

##Loading and preprocessing the data
Download the data file, read and store the content into the variable "con"

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

```{r,echo=FALSE}
library(data.table)
library(reshape2)
library(ggplot2)
library(mice)
library(lubridate)

url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
savefile<-"data.zip"
datafolder<-".\\data"
if(!file.exists(savefile))
{
    download.file(url,savefile)
    
    dir.create(datafolder)
    unzip(savefile,exdir = datafolder)
}

file<-dir(datafolder)
filepath<-paste(datafolder,file,sep="\\")
con<-fread(filepath)
```
##What is mean total number of steps taken per day?


```{r}
con2<- aggregate(steps~date,data = con,FUN=sum,na.rm=TRUE)
```
Histogram of the step number per day
```{r}
step<-con2$steps
#hist(step,breaks=9)
g<-ggplot(data=con2,aes(step))
p<-g+geom_histogram(breaks=seq(0, 22000, by=1500),col="red",fill="blue")
print(p)
```

Calculate the mean and median steps per day
```{r}
mean_step_con <- aggregate(steps~date,data = con,FUN=mean,na.rm=TRUE)
median_step_con <- aggregate(steps~date,data = con,FUN=median,na.rm=TRUE)
mean_median_step_con <- mean_step_con
mean_median_step_con$median_step <- median_step_con$steps
names(mean_median_step_con) <- c("date","mean_step","median_step")

knitr::kable(mean_median_step_con,caption="Table of mean and median step per day")

```

What is the average daily activity pattern?

```{r}
mean_daily_step_con<- aggregate(steps~interval,data = con,FUN=mean,na.rm=TRUE)
#mean_daily_step_con$interval<-sapply(mean_daily_step_con$interval, function(x) paste(substr(x,1,2),substr(x,3,4),sep=":"))

g<-ggplot(data=mean_daily_step_con,aes(interval,steps))
p<-g+geom_line(size=2)
print(p)
```



```{r}
max_interval<- which.max(mean_daily_step_con$steps)
max_step<-mean_daily_step_con$steps[max_interval]
```
###The interval contains max steps is `r max_interval`, and the average steps of that interval is `r max_step`


##Imputing missing values

Missing value of each interval on average
```{r}
table(con$date,is.na(con$steps))

```
###The number of NA row is  `r sum(is.na(con$steps))`

Use the interval average to impute NA
First reorder the the data frame con by the interval and calculate the mean by interval which outputs a variable as the same length as con but the all the steps values are the mean value
```{r}
con_reorder <- con[order(con$interval)]
mean_steps_by_interval<-tapply(con_reorder$steps,con_reorder$interval,function(x) mean(x,na.rm=T)*rep(1,length(x)))
```

Assign the mean value to replace "NA", which is achieved by addition
```{r}
NA_index <- is.na(con_reorder$steps)
con_reorder$steps[NA_index]<-0
con_reorder$steps<- con_reorder$steps + NA_index*unlist(mean_steps_by_interval)
```
Histogram of the steps after imputation

```{r}
g<-ggplot(data=con_reorder,aes(step))
p<-g+geom_histogram(breaks=seq(0, 22000, by=1500),col="red",fill="blue")
```

```{r}
mean_step_con <- aggregate(steps~date,data = con_reorder,FUN=mean,na.rm=TRUE)
median_step_con <- aggregate(steps~date,data = con_reorder,FUN=median,na.rm=TRUE)
mean_median_step_con <- mean_step_con
mean_median_step_con$median_step <- median_step_con$steps
names(mean_median_step_con) <- c("date","mean_step","median_step")

knitr::kable(mean_median_step_con,caption="Table of mean and median step per day after imputing")

```


##Are there differences in activity patterns between weekdays and weekends?

Create a new column "week_flag" indicating if the date is weekend or not

```{r}
week_index<-wday(ymd(con_reorder$date))>=6
week_flag_character <- character(ncol(con_reorder))
week_flag_character[week_index==0] <- "weekday"
week_flag_character[week_index==1] <- "weekend"

con_reorder$week_flag <- week_flag_character
```
```{r}
step_avg_weekend_weekday<-aggregate(steps~interval+week_flag,data=con_reorder,FUN=mean)
g<-ggplot(data=step_avg_weekend_weekday,aes(interval,steps))
p<-g+geom_line(size=2)+facet_wrap(~week_flag,scale="free")
print(p)
```
